---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# onnxr

<!-- badges: start -->
[![R-CMD-check](https://github.com/CoryMcCartan/onnxr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CoryMcCartan/onnxr/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/onnxr)](https://CRAN.R-project.org/package=onnxr)
<!-- badges: end -->

The **onnxr** package provides native R access to the [Open Neural Network
Exchange (ONNX) Runtime](https://onnxruntime.ai), which is a performant engine
for running machine learning models that are saved to a standardized format.
Rather than interfacing with ONNX via Python, as in the official
[onnx](https://cran.r-project.org/package=onnx) R package, this package directly
interfaces with the runtime's C++ API via [cpp11](https://cran.r-project.org/package=cpp11). This uses far less memory and does not require a Python and
TensorFlow installation.

Models saved to `.onnx` files can be loaded and run on various backends,
including CPUs and Apple's CoreML library.

## Installation

You can install **onnxr** from [CRAN](https://cran.r-project.org/package=onnxr) with:

``` r
install.packages("onnxr")
```

Or install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("CoryMcCartan/onnxr")
```

## Getting started

On first load of the package, you will need to call `onnx_install()` if
`libonnxruntime` is not already installed on your system.
This will download and install the latest version of the ONNX Runtime library, which is roughly 35 MB.

```{r eval=FALSE}
onnxr::onnx_install()
```

The package comes with an example `.onnx` GLM model for predicting *versicolor* vs *virginica* in the `iris` data.

```{r load}
library(onnxr)

model_path <- system.file("extdata", "glm_iris.onnx", package = "onnxr")
model <- onnx_model(model_path)
```

By default, the model will run on the CPU.
Other backends are available with the `backend` argument to `onnx_model()`.
For example, setting `onnx_model(model_path, backend = "coreml")` would load the model to run via Apple's CoreML library on Apple Silicon devices.

Printing the model object shows information on the input and output arrays.
```{r}
model
```

The model can be called by using `onnx_run()`, which accepts named or positional arguments matching the model's inputs.
```{r predict}
X = model.matrix(Species ~ 0 + ., data = iris)
output = onnx_run(model, X = X)
str(output)
```

See the vignettes for a more involved end-to-end example.

## Prior Art

This package is based off of the [nativeORT](https://cran.r-project.org/package=nativeORT) package by [Caleb Carr](https://github.com/calebmcarr).
