---
title: "End-to-end example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{End-to-end example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

To demonstrate the package in a more realistic setting, we'll show how to run an image detection network.
First, we'll download the YOLO26 (You Only Look Once) pretrained model from the Ultralytics repository.

```{r download}
model_url = "https://github.com/ultralytics/assets/releases/download/v8.4.0/yolo26n.onnx"
model_path = file.path(tempdir(), "yolo26n.onnx")
download.file(model_url, model_path, mode = "wb")
```

Then we can load the package and load the model.
```{r session}
library(nativeORT)

model = ort_model(model_path)
print(model)
```

We can see that the model takes in a single 640x640 RGB image stored as a 4-dimensional array.
The [model documentation](https://docs.ultralytics.com/guides/end2end-detection) explains that the output is a set of 300 bounding boxes, each with 6 values: the coordinates of the box (x1, y1, x2, y2), the confidence score, and the class index.
We can store the object types in a vector for later use.

```{r}
types = c(
    "person", "bicycle", "car", "motorcycle", "airplane", "bus", "train", 
    "truck", "boat", "traffic light", "fire hydrant", "stop sign", 
    "parking meter", "bench", "bird", "cat", "dog", "horse", "sheep", "cow", 
    "elephant", "bear", "zebra", "giraffe", "backpack", "umbrella", "handbag", 
    "tie", "suitcase", "frisbee", "skis", "snowboard", "sports ball", "kite", 
    "baseball bat", "baseball glove", "skateboard", "surfboard", "tennis racket", 
    "bottle", "wine glass", "cup", "fork", "knife", "spoon", "bowl", "banana", 
    "apple", "sandwich", "orange", "broccoli", "carrot", "hot dog", "pizza", 
    "donut", "cake", "chair", "couch", "potted plant", "bed", "dining table", 
    "toilet", "tv", "laptop", "mouse", "remote", "keyboard", "cell phone", 
    "microwave", "oven", "toaster", "sink", "refrigerator", "book", "clock", 
    "vase", "scissors", "teddy bear", "hair drier", "toothbrush"
)
```

Now we are ready to load an image (Vermeer's [*The Milkmaid*](https://en.wikipedia.org/wiki/The_Milkmaid_(Vermeer))) and run the model on it.

```{r load}
# load image from URL and center in 640x640 array
img_url = "https://uploads0.wikiart.org/images/johannes-vermeer/the-milkmaid.jpg!Large.jpg"
img = array(1, dim = model$input_shapes$images)
con = url(img_url, "rb")
raw = jpeg::readJPEG(readBin(con, "raw", n = 5e4))
close(con)
img[1, 1:3, 21:620, 53:587] = aperm(raw, c(3, 1, 2))

# helper to plot image
plot.image <- function(x) {
    par(mar = c(0, 0, 0, 0))
    rev_y = rev(1:dim(x)[3])
    x = 0.299 * x[1, 1, , ] + 0.587 * x[1, 2, , ] + 0.114 * x[1, 3, , ]
    image(
        t(x[rev_y, ]),
        col = gray.colors(256, start = 0, end = 1),
        asp = nrow(x) / ncol(x),
        axes = FALSE
    )
}
plot.image(img)
```

To run the model, we simply call `ort_run()` with the session and the input image.
The `simplify = TRUE` argument tells the function to return the single output array directly, instead of a named list.
We see that the output contains bounding box coordinates, confidence scores, and object type indices, as expected.

```{r run}
res = ort_run(model, img, simplify = TRUE)
dim(res)
head(res[1, ,])
```

Finally, we can pull out the bounding boxes with confidence scores above a certain threshold and plot them on top of the image.

```{r plot}
plot.image(img)
idx_conf = which(res[1, , 5] >= 0.2)
for (j in idx_conf) {
    x1 = res[1, j, 1] / 640
    y1 = 1 - res[1, j, 2] / 640
    x2 = res[1, j, 3] / 640
    y2 = 1 - res[1, j, 4] / 640
    rect(x1, y1, x2, y2, border = "#e00", lwd = 1)
    lbl = paste0(types[res[1, j, 6] + 1], " (", round(res[1, j, 5] * 100), "%)")
    text(x1, y1, lbl, adj = c(0, -0.5), cex = 0.8, col = "#800", font = 2)
}
```